### 2024-11-28 paper/paper.qmd
Draft upload

---
title: "Behind the Delays: Exploring Variations in Toronto’s Transit Disruptions"
subtitle: "A Comparative Analysis of Delay  Duration Across TTC Subway, Streetcar, and Bus Systems "
author: 
  - Claire Chang
thanks: "Code and data are available at: [https://github.com/ClaireUoft/Toronto_TTC_Transportation)."
date: today
date-format: long
abstract: "This paper uses a Bayesian Linear Regression Model to examine which Toronto Transit Commission (TTC) transit mode—subway, streetcar, or bus—experiences the highest duration of delays. We found that buses experience significantly longer delays compared to subways and streetcars. This finding is important because it highlights critical areas where transit improvements could alleviate the impact of delays on daily commuters. Overall, our study provides valuable insights into TTC transit reliability, informing efforts to enhance urban transportation efficiency and passenger experience."
format: 
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

# Load libraries
library(tidyverse)
library(arrow)
library(here)
library(ggplot2)
library(dplyr)
library(knitr)

library(kableExtra)
library(modelsummary)
library(lubridate)

# Load cleaned datasets
cleaned_combined_data <-read_parquet(here::here("data/02-analysis_data/cleaned_combined_data.parquet"))

# Load Model 1
model1 <- readRDS(here::here("models/first_model.rds"))
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

The remainder of this paper is structured as follows. @sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy. @sec-model outlines our modeling approach..., incorporating lessons learned from recent electoral cycles. Our predictions are under section of each model. @sec-discuss discusses the implications of our findings and suggests directions for future research. Finally, @sec-appendix evaluates methodology and survey copy.

# Data {#sec-data}

## Overview

In this project, we used data from the opendatatoronto dataset created by [@opendatatoronto]. This dataset provided bus, subway, streetcar records essential for our analysis. In this project, we used R[@citeR] and several R packages for data processing, analysis, and visualization. Specifically, tidyverse [@tidyverse], arrow[@arrow], here[@here], ggplot2[@ggplot2], while dplyr[@dplyr] was key for data manipulation tasks. For dynamic report generation, knitr[@knitr] and kableExtra [@kableExtra] used, providing enhanced formatting for outputs. Together, these packages enabled efficient data cleaning, analysis, and visualization throughout the study.

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

```{r}
#| label: tbl-cleaned-data-preview
#| tbl-cap: Preview of Cleaned Dataset for Bus, Subway, and Streetcar (One Example Each)
#| echo: false
#| warning: false
#| message: false

# Filter the dataset and ensure at least one row for each mode: bus, subway, streetcar
filtered_data <- cleaned_combined_data %>%
  filter(Transit_mode %in% c("Bus", "Subway", "Streetcar")) %>%
  group_by(Transit_mode) %>%
  slice_head(n = 1) # Get the first row for each Transit_mode group

# Display the filtered dataset
filtered_data |>
  kable() |>
  kable_styling(
    font_size = 6,
    latex_options = c("scale_down")
  ) |>
  column_spec(1, width = "2.5cm")

```

## Data Cleaning

## Outcome variables

### Delay Duration (Min Delay)

The main outcome variable in this study is the duration of delays, measured in minutes (Min Delay). This variable quantifies the length of each delay experienced by the TTC's different transit modes (subway, streetcar, bus). The primary goal of the analysis is to determine which transit mode exhibits the highest average delay duration.

To visualize the distribution of delay durations, a histogram @fig-histogram-delays is presented to provide an overview of the range of delay durations for each transit mode. This figure highlights the spread and common ranges of delay times, allowing us to observe whether there are particularly long delays for any given mode.

```{r}
#| label: fig-histogram-delays
#| fig-cap: "Histogram of Delay Duration (Filtered at 45 Minutes)"
#| echo: false
#| warning: false

# Filtering delays that are greater than 120 minutes
filtered_data <- cleaned_combined_data %>%
  filter(`Min Delay` <= 45)

# Plot histogram with the filtered data
ggplot(filtered_data, aes(x = `Min Delay`)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
  labs(
    title = "Histogram of Delay Duration",
    x = "Delay Duration (Minutes)",
    y = "Frequency"
  ) +
  theme_minimal()

```

Lastly, @fig-boxplot-delays displays a boxplot comparing delay durations across subway, streetcar, and bus transit modes. The boxplot offers insights into the median delays and potential outliers that may suggest unusual or extreme events, while showing variability in delay times across the different transit modes.

```{r}
#| label: fig-boxplot-delays
#| fig-cap: "Boxplot of Delay Duration by Transit Mode (Filtered by IQR)"
#| echo: false
#| warning: false

# Calculate IQR for Min Delay
delay_iqr <- cleaned_combined_data %>%
  summarise(
    Q1 = quantile(`Min Delay`, 0.25, na.rm = TRUE),
    Q3 = quantile(`Min Delay`, 0.75, na.rm = TRUE)
  )

iqr_value <- delay_iqr$Q3 - delay_iqr$Q1
upper_bound <- delay_iqr$Q3 + 1.5 * iqr_value

# Filter out delays above the upper bound
filtered_data <- cleaned_combined_data %>%
  filter(`Min Delay` <= upper_bound)

# Create the boxplot using filtered data
ggplot(filtered_data, aes(x = Transit_mode, y = `Min Delay`, fill = Transit_mode)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Delay Duration by Transit Mode (Filtered by IQR)",
    x = "Transit Mode",
    y = "Delay Duration (Minutes)"
  ) +
  theme_minimal()
```

## Predictor variables

### Transit Mode

The main predictor variable is Transit Mode, which includes three categories: subway, streetcar, and bus. Each transit mode has distinct characteristics, service areas, and operational challenges that may contribute to variations in delay duration.

To examine this relationship, @fig-transit plots the average delay duration for each transit mode. This chart helps us visualize which transit mode experiences the longest delays on average and illustrates the extent to which each mode contributes to the overall delay picture.

```{r}
#| label: fig-transit
#| fig-cap: "Average Delay Duration by Transit Mode"
#| echo: false
#| warning: false
#Bar Chart of Average Delay Duration by Transit Mode.

ggplot(cleaned_combined_data, aes(x = Transit_mode, fill = Transit_mode)) +
  geom_bar() +
  labs(
    title = "Frequency of Observations by Transit Mode",
    x = "Transit Mode",
    y = "Count"
  ) +
  theme_minimal()
```

### Time of Day

Another key predictor is Time of Day shown in @fig-time. This variable plays a crucial role in understanding how congestion and operational challenges fluctuate throughout the day. For example, delays might be more frequent and prolonged during peak hours (morning and evening rush periods) due to increased ridership.

```{r}
#| label: fig-time
#| fig-cap: "Average Delay by Time of Day"
#| echo: false
#| warning: false
# Line Graph for Average Delay by Time of Day
average_delay_time <- cleaned_combined_data %>%
  group_by(Time) %>%
  summarise(Average_Delay = mean(`Min Delay`, na.rm = TRUE))

ggplot(average_delay_time, aes(x = as.POSIXct(Time, format = "%H:%M"), y = Average_Delay)) +
  geom_line(color = "blue") +
  labs(
    title = "Average Delay Duration by Time of Day",
    x = "Time of Day",
    y = "Average Delay Duration (Minutes)"
  ) +
  theme_minimal()
```

@fig-time shows a line plot representing average delay duration across different hours of the day. This visualization allows us to determine the times at which each transit mode is most vulnerable to delays, and whether certain transit modes are disproportionately affected during rush hours.

### Day of the Week

```{r}
#| label: fig-day
#| fig-cap: "Delay Duration by Day of the Week"
#| echo: false
#| warning: false

# Calculate IQR and Filter Outliers
delay_iqr <- cleaned_combined_data %>%
  group_by(Day) %>%
  summarise(
    Q1 = quantile(`Min Delay`, 0.25, na.rm = TRUE),
    Q3 = quantile(`Min Delay`, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1
  )

# Join IQR data back to the main dataset
analysis_data_filtered <- cleaned_combined_data %>%
  left_join(delay_iqr, by = "Day") %>%
  filter(
    `Min Delay` >= (Q1 - 1.5 * IQR) & `Min Delay` <= (Q3 + 1.5 * IQR)
  )

# Boxplot of Delay Duration by Day of the Week without Outliers
figure5_filtered_boxplot <- analysis_data_filtered %>%
  ggplot(aes(x = Day, y = `Min Delay`, fill = Day)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Delay Duration by Day of the Week (Without Outliers)",
    x = "Day of the Week",
    y = "Delay Duration (minutes)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_x_discrete(limits = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

# Print the boxplot without outliers
print(figure5_filtered_boxplot)

```

The Day of the Week is an important predictor that may influence delays due to varying passenger demand and maintenance schedules. Weekdays typically see different patterns compared to weekends, with increased rider demand and therefore greater potential for delays.

@fig-day is a bar chart that shows the average delay duration for each day of the week, grouped by transit mode. It allows us to observe any weekly trends in delay duration, such as consistent weekday delays or weekend-specific issues, giving more insight into the operational aspects of the different modes. The predictor variables, including transit mode, time of day, and day of the week, help in assessing the specific conditions that affect the delay duration.

\\newpage

# Model {#sec-model}

The Bayesian linear regression model is designed to predict the duration of delays experienced by different TTC transit modes (subway, streetcar, or bus) based on various predictors such as transit mode, time of day, and day of the week. Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model Set-Up

Define $y_i$ as the duration of the delay (in minutes) for a given transit event. Let $\beta_i$ represent the transit mode (subway, streetcar, or bus), and $\gamma_i$ and $\delta_i$ represent the time and day of the week, respectively. The generative model is specified as follows:

\begin{align} 
y_i | \mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i + \delta_i \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\delta &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

Here:

\begin{itemize}
    \item $y_i$ is the observed delay duration for a specific transit event.
    \item $\alpha$ is the baseline average delay duration.
    \item $\beta_i$ captures the effect of transit mode (subway, streetcar, or bus).
    \item $\gamma_i$ captures the effect of the time of day (e.g., peak or off-peak).
    \item $\delta_i$ captures the effect of the day of the week.
    \item $\sigma$ is the standard deviation, representing unexplained variability in $y_i$.
\end{itemize}

This Bayesian model allows for estimating the contributions of transit mode, time, and day to delay duration while accounting for uncertainty in the data.

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

## Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

## Assumptions and Limitations

# Results

These statistics indicate that, on average, TTC streetcars experienced a delay duration of approximately 15.7 minutes, while buses faced an average delay of 12.5 minutes, and subways experienced the shortest average delay of 9.3 minutes. The total number of delay events recorded was 50,039 for buses, 21,979 for subways, and 12,107 for streetcars.

```{r}
#| echo: false
#| eval: true
#| label: tbl-summary-stats
#| warning: false
# Summary Statistics for Delay Events by Transit Mode

# Calculate the average delay duration and count for each transit mode
transit_mode_summary <- cleaned_combined_data %>%
  group_by(Transit_mode) %>%
  summarise(
    `Average Delay Duration (Minutes)` = mean(`Min Delay`, na.rm = TRUE),
    `Count of Events` = n()
  ) %>%
  arrange(desc(`Count of Events`))

# Create the summary statistics table
kable(transit_mode_summary, caption = "Summary statistics of delay events for TTC transit modes") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)

```

The bar plot visually confirms that streetcars have consistently higher average delay durations compared to buses and subways, aligning with the numerical analysis."

```{r}
#| echo: false
#| warning: false
#| message: false

cleaned_combined_data %>%
  group_by(Transit_mode) %>%
  summarise(Average_Delay = mean(`Min Delay`, na.rm = TRUE)) %>%
  ggplot(aes(x = Transit_mode, y = Average_Delay, fill = Transit_mode)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Delay Duration by Transit Mode", y = "Average Delay (Minutes)", x = "Transit Mode") +
  theme_minimal()

```

The model summary @tbl-modelresults results indicate that, compared to buses, streetcars and subways have shorter delays, with average reductions of 6.8 and 18.8 minutes, respectively. Delays also vary by day of the week, with Tuesday and Wednesday showing slightly shorter delays. However, the model explains only about 3% of the variance in delay duration (R² = 0.033), and the RMSE of 44.88 suggests moderate prediction error. While transit mode and day of the week have some influence, much of the variation in delay duration remains unexplained.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Summary of model results evaluating the relationship between transit mode and delay duration"
#| warning: false

modelsummary::modelsummary(
  list(
    "model1" = model1
  )
)
```

The @fig-predict_delay graph indicates that bus face the highest average delay durations, peaking at over 40 minutes in the early morning. Streecar have moderate delays throughout the day, averaging between 15-20 minutes, while subways consistently experience the lowest delays, averaging below 10 minutes. The significant early morning peak for streetcars suggests this as a critical period for intervention to reduce delays.

```{r}
#| label: fig-predict_delay
#| fig-cap: Predicting average delay duration by time of day for each transit mode
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Predicting average delay duration by time of day for each transit mode
# Create a `Time_of_Day` column by extracting the hour from `Time`
cleaned_combined_data <- cleaned_combined_data %>%
  mutate(Time_of_Day = hour(hms(Time)))

# Calculate average delay duration by time of day for each transit mode
avg_delay_time_of_day <- cleaned_combined_data %>%
  group_by(Time_of_Day, Transit_mode) %>%
  summarise(Average_Delay = mean(`Min Delay`, na.rm = TRUE)) %>%
  ungroup()

# Plotting average delay by time of day for each transit mode
ggplot(avg_delay_time_of_day, aes(x = Time_of_Day, y = Average_Delay, color = Transit_mode)) +
  geom_line(size = 1.2) +
  labs(title = "Predicted Average Delay Duration by Time of Day for Each Transit Mode",
       x = "Time of Day (Hourly)",
       y = "Average Delay Duration (Minutes)",
       color = "Transit Mode") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Discussion {#sec-discuss}

Some questions that a good discussion would cover include (each of these would be a sub-section of something like half a page to a page): What is done in this paper? What is something that we learn about the world? What is another thing that we learn about the world? What are some weaknesses of what was done? What is left to learn or how should we proceed in the future?

## First discussion point {#sec-first-point}

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Limitations and next steps

\newpage

\appendix

# Appendix {#sec-appendix}

Surveys, sampling, and observational data appendix Please include an appendix where you focus on some aspect of surveys, sampling or observational data, related to your paper. This should be an in-depth exploration, akin to the "idealized methodology/survey/pollster methodology" sections of Paper 2. Some aspect of this is likely covered in the Measurement sub-section of your Data section, but this would be much more detailed, and might include aspects like simulation and linkages to the literature, among other aspects.

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

@fig-stanareyouokay-1 is a trace Plot that shows the sampled values of each parameter from the MCMC chains over the iterations. Ideally, these chains should overlap well and display consistent movement across the same regions. From the trace plot, the chains for all parameters appear to have converged and are overlapping consistently, without any clear trends or divergences. This indicates good mixing and suggests the MCMC algorithm has likely converged.

@fig-stanareyouokay-2 is a Rhat Plot that shows the potential scale reduction factor, which assesses convergence of the MCMC chains. The Rhat statistic should be close to 1, typically below 1.1, to indicate good convergence. In this plot, the Rhat values for all parameters are very close to 1, suggesting the MCMC algorithm has converged, and there is no significant difference between the variances within and across chains.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(model1, "trace")

plot(model1, "rhat")
```

\newpage

# References